An Infinite de Bruijn Sequence

### 2016-12-25 infdebruijn.tex
With a proof that the firsts are ordered lexicographically

A key word is a word of length $n$ that is maximal in right-to-left (Arabic) lexicographic order among its cyclic rotations. Let $kw_0,kw_1,kw_2,\dots$ be an enumeration of all key-words, ordered in right-to-left (Arabic) lexicographic order. Let $C_m=\{R^i(kw_m): i=1,\dots,n\}$ be the cycle of $kw_m$. We order the elements of $C_m$ as follows:  if $kw_m=0^l(\sigma+1)w$, then $w0^l(\sigma+1)$ is the first word in $C_m$, and each word $w'$ is followed by $R(w')$. The last word in $C_m$ is $(\sigma+1)w0^l$.\\

@@

\begin{theorem}
	For a  word $\sigma w$ in $D$, let $nxt(\sigma w)$ be the successor of $\sigma w$ in $D$. Then,
	$$nxt(\sigma w)=\left\{ \begin{array}{lr}
	w(\sigma+1), & last((\sigma+1)w)\\
	w0, & last(\sigma w)\\
	w\sigma, & otherwise
	\end{array}\right.$$
\end{theorem}

@@

\begin{definition}
	We say that $C_m$ is embedded in $C_k$ if $first(C_k)<first(C_m)\leq last(C_m)<last(C_k)$.\\ 
	In addition, $C_m$ is said to be immediately embedded in $C_k$ if there is no $C_l$ such that $C_m$ is embedded in $C_l$ and $C_l$ is embedded in $C_k$.\\ 
	We define by inductively the statement: ``$C_m$ is $r$-embedded in $C_k$":
	\begin{itemize}
		\item $C_m$ is $1$-embedded in $C_k$ if it is immediately embedded in $C_k$.
		\item $C_m$ is $r$-embedded in $C_k$ if there is a cycle $C_l$ such that $C_m$ is $r-1$-embedded in $C_l$ and $C_l$ is immediately embedded in $C_k$. 
	\end{itemize} 
\end{definition}

@@

	As a result, we have:
	$$nxt(last(C_{m+r}))=nxt((\sigma+1+r)0^jw0^i)\in\{\ 0^jw0^i0, 0^jw0^i(\sigma+2+r)\}.$$ If $nxt((\sigma+1+r)0^jw0^i)= 0^jw0^i(\sigma+2+r)$, then $last((\sigma+2+r)0^jw0^i)$ which implies $first(0^jw0^i(\sigma+2+r))$. But $first(last(C_{m+r}))$ contradicts the maximality of our sequence of cycles. Hence, $nxt((\sigma+1+r)0^jw0^i)= 0^jw0^i0\in C_k$. Now, $0^{i+1+j}w$ is a rotation of $0^jw0^i0$ thus $0^{i+1+j}w\in C_k$ as required.

@@

\begin{lemma}
	\label{first-ordering}
	For every $m\in\N$, $first(C_m)<first (C_{m+1})$.
\end{lemma}

\begin{proof}
	TBD
\end{proof}

>>>>

A key word is an $n$-length word that is (arabic) maximal among its rotations. Let $kw_0,kw_1,kw_2,\dots$ be an enumeration of all key-words, ordered lexicographically. Let $C_m$ be the cycle of $kw_m$. We order the elements of $C_m$ as follows:  if $kw_m=0^l(\sigma+1)w$, then $w0^l(\sigma+1)$ is the first sequence in $C_m$, and each word $w'$ is followed by $R(w')$. The last word in $C_m$ is $(\sigma+1)w0^l$.\\

@@

\begin{theorem}
	For a  word $\sigma w$ in $D$, let $nxt(\sigma w)$ be the successor of $\sigma w$ in $D$. Then,
	$$nxt(\sigma w)=\left\{ \begin{array}{ll}
	w(\sigma+1) & \text{if }last((\sigma+1)w);\\
	w0          & \text{if } last(\sigma w);\\
	w\sigma     & \text{otherwise}
	\end{array}\right.$$
\end{theorem}

@@

\begin{definition}
	We say that $C_m$ is embedded in $C_k$, if $C_m=C_k$ or $first(C_k)<first(C_m)\leq last(C_m)<last(C_k)$.\\ 
	In addition, $C_m$ is said to be immediately embedded in $C_k$ if there is no $C_l$ such that $C_m$ is embedded in $C_l$ and $C_l$ is embedded in $C_k$.\\ 
	We define by inductively the statement: ``$C_m$ is $r$-embedded in $C_k$":
	\begin{itemize}
		\item $C_m$ is $0$-embedded in $C_k$ if $C_m=C_k$.
		\item $C_m$ is $r$-embedded in $C_k$ if there is a cycle $C_l$ such that $C_m$ is $r-1$-embedded in $C_l$ and $C_l$ is immediately embedded in $C_k$. 
	\end{itemize} 
\end{definition}

@@

	As a result, we have:
	$$nxt(last(C_{m+r}))=nxt((\sigma+1+r)0^jw0^i)\in\{\ 0^jw0^i0, 0^jw0^i(\sigma+2+r)\}.$$ If $nxt((\sigma+1+r)0^jw0^i)= 0^jw0^i(\sigma+2+r)$, then $last((\sigma+2+r)0^jw0^i)$ which implies $first(0^jw0^i(\sigma+2+r))$. But $first(nxt(last(C_{m+r})))$ contradicts the maximality of our sequence of cycles. Hence, $nxt((\sigma+1+r)0^jw0^i)= 0^jw0^i0\in C_k$. Now, $0^{i+1+j}w$ is a rotation of $0^jw0^i0$ thus $0^{i+1+j}w\in C_k$ as required.

@@

\begin{lemma}
	\label{both-embedded}
	Assume that $last(C_k)<first(C_m)$ and $C_k,C_m$ are both immediately embedded in a cycle $C$. Write $key(C_k)=0^lw$ where $w$ starts with a non-zero letter. Then, $key(C_m)=0^lw'$ where $w<_{lex}w'$.
\end{lemma}
\begin{proof}
	First, if there is a sequence of cycles from $C_k$ and $C_m$, the claim follows from Lemma \ref{sequence}. Otherwise, there is some $u\in C$ such that
	$$last(C_k)<u<first(C_m).$$
	Consider a  maximal sequence of cycles that starts with $C_k$. This sequence ends in some cycle $C_{k'}$. Similarly, consider a maximal sequence of cycles that ends in $C_m$ and let $C_{m'}$ be the first element in this sequence. Therefore,
	$$last(C_{k'}) <v<first(C_{m'})\Longrightarrow v\in C.$$

	Write $key(C_k)=0^l(\sigma+1)w_1$ (namely, $w=(\sigma+1)w_1$) and write  $w_1=0^jw_2$ where $w_2$ starts with non-zero letter. Hence, 
	\begin{equation}
	\label{ck}
	key(C_k)=0^l(\sigma+1)0^jw_2.
	\end{equation}
	By Lemma \ref{sequence}, $key(C_{k'})=0^l(\sigma+1+r)0^jw_2$, and hence $last(C_{k'})=(\sigma+1+r)0^jw_20^l$. Let $u_1$ be the successor of $last(C_{k'})$. Thus,
	$$u_1=0^jw_20^{l+1}.$$
	Now, let $u_2$ be the predecessor of $first(C_{m'})$. Thus,
	$$u_2=0^{j_2}w_20^{l+1+j_1}, \text{ where } j_1+j_2=j.$$
	In addition, since $C_{m'}$ is embedded in $C$, $u_2\neq last(C)$ thus $j_2>0$. As a result, $$first(C_{m'})=0^{j_2-1}w_20^{l+1+j_1}1.$$
	We get that $key(C_{m'})=0^{l+1+j_1}10^{j_2-1}w_2$. Hence, by Lemma \ref{sequence}, 
	\begin{equation}
	\label{cm}
	key(C_m)=0^{l+1+j_1}(1+t)0^{j_2-1}w_2.
	\end{equation}
	By Equation \ref{ck}, $w=(\sigma+1)0^{j}w_2$. By Equation \ref{cm}, $w'=0^{1+j_1}(1+t)0^{j_2-1}w_2$. We see that indeed $w<_{lex}w'$.
\end{proof}

\begin{lemma}
	\label{first-ordering}
	If $k<m$, then $first(C_k)<first (C_m)$.
\end{lemma}

\begin{proof}
	Since $<$ is a linear ordering, we can prove an equivalent statement: $$first(C_k)<first(C_m) \Longrightarrow k<m.$$ We take such cycles $C_k$ and $C_m$. By the parentheses property, either $C_m$ is embedded in $C_k$, or $C_m$ is entirely after $C_k$. If $C_m$ is embedded in $C_k$, by Corollary \ref{embedding-deletes-zeroes} we get that $k<m$. It is left to deal with the case that $last(C_k)<first(C_m)$. We consider two cases.
	\begin{description}
		\item[Case 1.] $C_k$ and $C_m$ are both embedded in some cycle $C$.

		In this case, we can find cycles $C_{k'}$ and $C_{m'}$ such that
		\begin{enumerate}
			\item $C_k$ is embedded in $C_{k'}$ and $C_{m}$ is embedded in $C_{m'}$

			\item $C_{k'}$ and $C_{m'}$ are immediately embedded in $C$.

		\end{enumerate}

		By item 1 and Lemma~\ref{embedding-deletes-zeroes}, we can write $key(C_k)=uv$ and $key(C_{k'})=0^{|u|}v$. Write 
		$$v=0^lv_1$$ 
		where $v_1$ starts with a non-zero letter thus $key(C_{k'})=0^{|u|+l}v_1$. By item~2 and Lemma~\ref{both-embedded}, $key(C_{m'})=0^{|u|+l}v_2$ where 
		$$v_1<_{lex}v_2.$$

		Write $v_2=0^r v_2'$ where $v_2'$ starts with a non zero letter. Hence, $$v_1=xv_1'$$ 
		where $|x|=r$, $|v_1'|=|v_2'|$, and 
		$$v_1'<_{lex} v_2'.$$
		Since $key(C_{m'})=0^{|u|+l}v_2 =0^{|u|+l+r}v_2'$, by item~2 and Lemma \ref{embedding-deletes-zeroes}, $key(C_m)=u'v_2'$. Recall that $key(C_k)=uv=u0^lv_1=u0^lxv_1'$. Since $v_1'<_{lex}v_2'$, $key(C_k)<_{lex}key(C_m)$ thus $k<m$ as required.

		\item[case 2.] There is no cycle $C$ such that $C_k$ and $C_m$ are embedded in $C$.

		Take a cycle $C_{k'}$ such that $C_k$ is $r_1$-embedded in $C_{k'}$ and $r_1$ is maximal with respect to this property. Similarly, take a cycle $C_{m'}$ such that $C_{m}$ is $r_2$-embedded in $C_{m'}$ and $r_2$ is maximal as possible. By the parentheses property, $C_{k'}$ is entirely before $C_{m'}$. $C_{k'}$ and $C_{m'}$ are not embedded in any cycle so there is a sequence of cycles from $C_{k'}$ to $C_{m'}$. Write $key(C_k)=u(\sigma+1)v$ such that $key(C_{k'})=0^{|u|}(\sigma+1)v$ (by Lemma \ref{embedding-deletes-zeroes}). Hence, $key(C_{m'})=0^{|u|}(\sigma+1+r)v$ where $r>1$ (by Lemma \ref{sequence}). Finally, by Lemma \ref{embedding-deletes-zeroes} we get $key(C_m)=u'(\sigma+1+r)v$. We see that $key(C_k)=u(\sigma+1)v<_{lex}u'(\sigma+1+r)v=key(C_m)$. Therefore, $k<m$.

	\end{description}
\end{proof}
